Data Prep: IUCN and AquaMaps

### 2016-02-01 data_explore/data_prep.Rmd
updating species list (spp_all.csv) with missed or dropped species

@@

@@

spp_am_v_iucn <- spp_all %>%
  filter(str_detect(spatial_source, 'iucn') & !is.na(am_sid)) %>%
  unique()
write.csv(spp_am_v_iucn, file.path(dir_data, 'spp_am_v_iucn.csv'), 
          row.names = FALSE)

### pull spp_iucn_marine.csv from git-annex and bring to local repo
spp_iucn_marine <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_marine.csv'), 
                            stringsAsFactors = FALSE)
spp_iucn_marine <- spp_iucn_marine %>%
  dplyr::select(sciname, iucn_sid, iucn_category = category, 
                popn_trend, parent_sid, subpop_sid) %>%
  unique()
write.csv(spp_iucn_marine, file.path(dir_data, 'spp_iucn_marine.csv'), 
          row.names = FALSE)

>>>>

``` {r pull_files_to_local, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### pull spp_iucn_marine.csv from git-annex and bring to local repo
spp_iucn_marine <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_marine.csv'), 
                            stringsAsFactors = FALSE)
spp_iucn_marine <- spp_iucn_marine %>%
  dplyr::select(sciname, iucn_sid, iucn_category = category, 
                popn_trend, parent_sid, subpop_sid) %>%
  unique()
write.csv(spp_iucn_marine, file.path(dir_data, 'spp_iucn_marine.csv'), 
          row.names = FALSE)

@@

### But, some of the extracted IUCN maps don't appear on this list! add 'em in.
spp_iucn_marine   <- read.csv(file.path(dir_data, 'spp_iucn_marine.csv'), 
                              stringsAsFactors = FALSE)
spp_iucn_maps_new <- read.csv(file.path(dir_data, 'spp_iucn_maps.csv'), 
                              stringsAsFactors = FALSE) %>%
  filter(!sciname %in% spp_all$sciname) %>%
  filter(!str_detect(sciname, 'subpop')) %>% ### a Hawaiian subpop... ditch it!
  select(-iucn_subpop)
spp_all1 <- spp_all %>%
  full_join(spp_iucn_maps_new)

### Also, some of the AM listings have disappeared due to name conflicts
### - aliases, etc.
spp_am_new <- read.csv(file.path(dir_data, 'spp_am.csv'), 
                       stringsAsFactors = FALSE) %>%
  rename(am_sid = speciesid) %>%
  filter(!am_sid %in% spp_all$am_sid) %>%
  dplyr::select(am_sid, sciname, reviewed, iucn_sid = iucn_id,
                popn_category = iucn_code) %>%
  mutate(spatial_source = 'am',
         iucn_sid      = as.integer(iucn_sid),
         popn_category = ifelse(popn_category == 'LR/lc', 'LC', popn_category),
         popn_category = ifelse(popn_category == 'LR/nt', 'NT', popn_category),
         popn_category = ifelse(popn_category == 'N.E.',  'NE', popn_category))

### 600 new spp_am not in spp_all from remote source, by am_sid.
### 12 of these, however, list iucn_sids that match existing iucn_sid.
### * for species with iucn_sid that match, and no am_sid, add am_sid to 
###   existing record (use existing values for other fields; drop new values).
### * for species with iucn_sid that match, and with a different am_sid, no action.
### * for species with non-matching or NA iucn_sid, just add to list.
spp_am_new_iucnmatch <- spp_am_new %>% 
  filter(!is.na(iucn_sid) & iucn_sid %in% spp_all$iucn_sid) %>%
  dplyr::select(iucn_sid, am_sid_new = am_sid, reviewed_new = reviewed)
spp_all2 <- spp_all1 %>%
  left_join(spp_am_new_iucnmatch, by = 'iucn_sid') %>%
  mutate(am_sid   = ifelse(is.na(am_sid) & !is.na(am_sid_new), am_sid_new, am_sid),
         reviewed = ifelse(is.na(am_sid) & !is.na(am_sid_new), reviewed_new, reviewed)) %>%
  select(-am_sid_new, -reviewed_new)
spp_all <- spp_all2 %>%
  full_join(spp_am_new %>%
              filter(!iucn_sid %in% spp_am_new_iucnmatch$iucn_sid),
            by = c("am_sid", "sciname", "reviewed", "iucn_sid", "popn_category", "spatial_source"))

@@

spp_am_v_iucn <- spp_all %>%
  filter(str_detect(spatial_source, 'iucn') & !is.na(am_sid)) %>%
  unique()
write.csv(spp_am_v_iucn, file.path(dir_data, 'spp_am_v_iucn.csv'), 
          row.names = FALSE)
```

``` {r pull_files_to_local, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### pull spp_iucn_marine.csv from git-annex and bring to local repo
spp_iucn_marine <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_marine.csv'), 
                            stringsAsFactors = FALSE)
spp_iucn_marine <- spp_iucn_marine %>%
  dplyr::select(sciname, iucn_sid, iucn_category = category, 
                popn_trend, parent_sid, subpop_sid) %>%
  unique()
write.csv(spp_iucn_marine, file.path(dir_data, 'spp_iucn_marine.csv'), 
          row.names = FALSE)
